Algebraic Tori in Unitary Groups

### 2014-07-02 Paper.tex
Explain role of "\lambda" for algebras with involution
Explain analog of importance of lambda in Tr(\lambda x\sigma(y)) for
algebras with involution

If in the above $B_1$ and $B_2$ are Hermitian then $\sigma(\lambda)=\lambda$.
\end{theorem}
\begin{proof}
We have the following calculation:
\[ B_2(\lambda y,x) = B_1(y,x) = B_1(x,y) = B_2(\lambda x,y) = B_2(x,\sigma(\lambda) y) = B_2(\sigma(\lambda)y,x) \]
Alternatively:
\[ B_2(\lambda y,x) = B_1(y,x) = \overline{B_1(x,y)} = B_2(\lambda x,y) = \overline{B_2(x,\sigma(\lambda) y)} = B_2(\sigma(\lambda)y,x) \]
And this holds for all $x,y$.
Hence, by the uniqueness of $\lambda$ from above, $\lambda=\sigma(\lambda)$.
\end{proof}
\begin{corollary}
All the non-degenerate symmetric bilinear forms on $E$ are of the form:
\[ Tr_{E/K}(\lambda x\sigma(y) \]
All the non-degenerate Hermitan forms on $E$ are of the form:
\[ Tr_{E/A}(\lambda x\sigma(y) \]
where $\lambda \in E^\sigma$ is a unit.
In both cases, the isomorphism class of the form depends only on $\lambda \in (E^\sigma)^\times/N_{E/E^\sigma}(E^\times)$.
\end{corollary}

>>>>

If in the above $B_1$ and $B_2$ are Hermitian then $\sigma(\lambda)=\lambda$.
\end{theorem}
\begin{proof}
We have the following calculation:
\[ B_2(\lambda y,x) = B_1(y,x) = B_1(x,y) = B_2(\lambda x,y) = B_2(x,\sigma(\lambda) y) = B_2(\sigma(\lambda)y,x) \]
Alternatively:
\[ B_2(\lambda y,x) = B_1(y,x) = \overline{B_1(x,y)} = B_2(\lambda x,y) = \overline{B_2(x,\sigma(\lambda) y)} = B_2(\sigma(\lambda)y,x) \]
And this holds for all $x,y$.
Hence, by the uniqueness of $\lambda$ from above, $\lambda=\sigma(\lambda)$.
\end{proof}
\begin{corollary}
All the non-degenerate symmetric bilinear forms on $E$ are of the form:
\[ Tr_{E/K}(\lambda x\sigma(y)) \]
All the non-degenerate Hermitan forms on $E$ are of the form:
\[ Tr_{E/A}(\lambda x\sigma(y)) \]
where $\lambda \in E^\sigma$ is a unit.
In both cases, the isomorphism class of the form depends only on $\lambda \in (E^\sigma)^\times/N_{E/E^\sigma}(E^\times)$.
\end{corollary}

\subsubsection{Algebra's with involutions}

\begin{theorem}
Let $A$ be a central simple algebra over $K$.
Let $E$ be an \'etale algebra over $k$ which embeds into $A$ as a maximal \'etale subalgebra.
Suppose $\tau_1,\tau_2$ are two involutions on $A$ which preserve $E$ and such that $\tau_1|_E = \tau_2|_E$ then:
\[ \tau_1 = {\rm Int}_\lambda\circ\tau_2 \]
for some $\lambda \in E^\times$.
\end{theorem}
\begin{proof}
Via the classification of involutions on $A$
\TODO{cite book of involutions or above}
 we know $\tau_1 = {\rm Int}_\lambda\circ\tau_2$ for $\lambda \in A^\times$.

However, since $\tau_1$ and $\tau_2$ both preserve $E$ it follows that ${\rm Int}_\lambda$ must preserve $E$ and thus at least normalize $E$.
Moreover, as $\tau_1|_E = \tau_2|_E$ we find that ${\rm Int}_\lambda$ acts trivially on $E$ and thus $\lambda$ centralizes $E$. We conclude that $\lambda\in E$ by checking over the algebraic closure.
\end{proof}

\begin{theorem}
If the involutions $\tau_1$ and $\tau_2$ are both orthogonal or both Hermitian then $\sigma(\lambda) = \lambda$.

If one of $\tau_1$ is simplectic, the other orthogonal, then $\sigma(\lambda) = -\lambda$.
\end{theorem}

\begin{corollary}
Fixing any one involution $\tau$ on $A$ (either orthogonal or Hermitian) preserving $E$.
All of the involutions on $A$ preserving $E$ are of the form:
\[ {\rm Int}_\lambda\circ\tau \]
where $\lambda \in (E^\sigma)\times$.
\end{corollary}

\begin{theorem}
Let $A$ be a central simple algebra over $K$.
Let $E$ be an \'etale algebra over $k$ which embeds into $A$ as a maximal \'etale subalgebra.
Let $\rho_1,\rho_2 : E\hookrightarrow A$ be two embeddings.

There exists $\gamma \in A^\times$ such that $\rho_1 = {\rm Int}_\gamma\rho_2$.
\end{theorem}
\begin{proof}[Sketch]
\TODO{Complete the sketch}

Let $g\in A \otimes_k \overline{k}$ realize ${\rm Int}_g\rho_2$.

"It is well known" that this exists over the algebraic closure.
  - It is well known that we can conjugate so that the images are equal, that we can obtain the isomorphism on the nose is the fact that weil group is big.

We can verify then that $g\sigma(g^{-1})$ gives an element of $E^\times$.
Thus $g\sigma(g^{-1}) = e\sigma(e^{-1})$ for $e\in E^\times$ (by Hilbert's 90) so that $ge^{-1}$ is a galois stable element realizing the isomorphism.
\end{proof}

\begin{remark}
The importance of the above theorem is that it allows us to interchange the following two problems:
\begin{enumerate}
\item Understand rational conjugacy classes of maps $\rho: (E,\sigma) \hookrightarrow (A,\tau)$
\item Understand the rational conjugacy classes of $\tau$ on $A$ extending $\sigma$ on $E$ with respect to a fixed embedding $E\hookrightarrow A$.
\end{enumerate}
By the above corralary \ref{TODO} this amounts to a thorough understanding of the rational conjugacy classes of the involutions $\tau_\lambda =  {\rm Int}_\lambda\circ\tau$.

\TODO{at some point need to explain why $\tau_\lambda$ and $\tau$ are not rationally conjugate}
Key point is that $\tau(\lambda x \lambda^{-1}) = \tau(\lambda^{-1}) \tau(x) \tau(\lambda)$ so the involutions $\tau$ and ${\rm Int}_{\tau(\lambda)}\circ \tau\circ {\rm Int}_\lambda$ are rationally conjugate.
\end{remark}
